---
title: "300-2017-09-17-example-containing-start-instead-of-begin"
author: ''
output:
  html_document:
    css: ~/Documents/headings.css
    fig_caption: yes
    highlight: default
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    pandoc_args: --variable=geometry:margin=0.75in
    toc: no
fontsize: 10pt
---
<A NAME="top"> </A>

```{r global_options, include=FALSE}
require(knitr)
opts_chunk$set(eval = TRUE, echo = TRUE, fig.height = 8, fig.width = 8, tidy=FALSE)
```

`r getwd()`  

Last knit was `r Sys.time()`

Copied from "209-2017-09-12-example-containing-exclusion_region.Rmd"


# A little history

The two data sets in `gatars`:

`alternative_example` and `null_example` I will be creating here.  

There is some history.  From `lacie/aaa/packages/gritsr2/gails-stuff/a-chronicle/04-v-0.1.4.html` I see that gritsr2 contains the following four data sets
```
bim_original.RData
exclusion_region.RData
genotype_bim_original.RData
hotspot.RData
```
which I moved from `d-2016-04-20/06-adding-valeries-data/gritsr_packages_RData`

These files were created by “107-2016-10-24-description-adding-valeries-data.Rmd”

The code below, I copied from "lacie/aaa/packages/gritsr2/gails-stuff/b-examples/03-v-0.1.4/a-build-hg38/201-gritsr-example-dot-RData"



```{r}
library(gritsr2)
set.seed(1)
genotype_bim_original = genotype_bim_original[1:5, ]
params_example = params_example_fn(
  bim_original, genotype_bim_original, hotspot,
  NNN = 200,
  p_subset_genotype = .2,
  beta_causal_snps = .85,
  causal_snps = 3:5)
alternative_example = example_fn(params_example)
alternative_example$params_example = params_example
save(alternative_example, file = "alternative_example.RData")
```


```{r}
bim = alternative_example$bim
genotype = alternative_example$genotype
fam = alternative_example$fam
target_markers = alternative_example$target_markers
Psi = alternative_example$Psi
```

```{r echo = TRUE}
str(bim)
str(genotype)
str(fam)
str(target_markers)
str(Psi)
```

# `null_example`

```{r}
set.seed(1)
genotype_bim_original = genotype_bim_original[1:5, ]
params_example = params_example_fn(
  bim_original, genotype_bim_original, hotspot,
  NNN = 200,
  p_subset_genotype = .2,
  beta_causal_snps = 0,
  causal_snps = 3:5)
null_example = example_fn(params_example)
null_example$params_example = params_example
save(null_example, file = "null_example.RData")
```


```{r}
bim = null_example$bim
genotype = null_example$genotype
fam = null_example$fam
target_markers = null_example$target_markers
Psi = null_example$Psi
```

```{r echo = TRUE}
str(bim)
str(genotype)
str(fam)
str(target_markers)
str(Psi)
```
