---
title: "305-2017-11-08-example-after-basic_and_optimized_lu_fn"
author: ''
output:
  html_document:
    css: ~/Documents/headings.css
    fig_caption: yes
    highlight: default
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    pandoc_args: --variable=geometry:margin=0.75in
    toc: no
fontsize: 10pt
---
<A NAME="top"> </A>

```{r global_options, include=FALSE}
require(knitr)
opts_chunk$set(eval = TRUE, echo = TRUE, fig.height = 8, fig.width = 8, tidy=FALSE)
```

`r getwd()`  

Last knit was `r Sys.time()`


# A little history for the 305 file

The data set in `gatars`: `gatars_example`  I will be creating here.  

There is some history.  From `lacie/aaa/packages/gritsr2/gails-stuff/a-chronicle/04-v-0.1.4.html` I see that gritsr2 contains the following four data sets
```
bim_original.RData
exclusion_region.RData
genotype_bim_original.RData
hotspot.RData
```

`bim_original` is a data.frame containing the columns `chromosome`, `snp`, and `bp` and 124303 rows, so these are possibilities for sampling sets.


`genotype_bim_original` is a data.frame containing the columns `chromosome`, `snp`, and `bp` and 47 rows, so these are possibilities for target markers.


`exclusion_region` is a data.frame containing the columns `chromosome` `begin` and `end` and 113 rows.

`hotspot` is a data.frame with 25657 rows and the columns `chromosome`, `center`, `start` and `end`.



which I moved from `d-2016-04-20/06-adding-valeries-data/gritsr_packages_RData`

These files were created by “107-2016-10-24-description-adding-valeries-data.Rmd”

The code below, I copied from "lacie/aaa/packages/gritsr2/gails-stuff/b-examples/03-v-0.1.4/a-build-hg38/201-gritsr-example-dot-RData"


```{r}
library(gatars)
library(gritsr2)
set.seed(2)
RData_file = "gatars_example.RData"
source_0 = getwd()
source_1 = strsplit(source_0, "/")[[1]]
source = paste0(paste(source_1[(length(source_1) - 2):length(source_1)], collapse = "/"),
      "/305-2017-11-08-example-after-basic_and_optimized_lu_fn.Rmd")
genotype_bim_original = genotype_bim_original[1:5, ]
params_example = params_example_fn(
  bim_original, genotype_bim_original, hotspot,
  NNN = 200,
  p_subset_genotype = .2,
  beta_causal_snps = .85,
  causal_snps = 3:5)
gatars_example = example_fn(params_example)
names(gatars_example)
str(gatars_example$bim)
str(gatars_example$exclusion_region)
str(gatars_example$genotype)

str(gatars_example$phenotype)
str(gatars_example$Psi)
gatars_example$target_markers
gatars_example$Psi[1:10, 1:10]
gatars_example$params_example = params_example

additional_attributes = list(
  RData_file = RData_file,
  when_created = Sys.time(),
  source = source
)
attributes(gatars_example) = c(attributes(gatars_example),
                                    additional_attributes)
                                               
save(gatars_example, file = RData_file)
```

```{r}
# here i am calling `gatars`
bim = gatars_example$bim
exclusion_region = gatars_example$exclusion_region
genotype = gatars_example$genotype
phenotype = gatars_example$phenotype
Psi = gatars_example$Psi
target_markers = gatars_example$target_markers[3:5]
NNN = nrow(phenotype)
first_ten = 1:10
last_ten = NNN - (9:0)

# Creating the sampling_set
set.seed(42)
epsilon = 0.01
exclusion_region = NULL
sampling_set = gatars_sampling_set(
  bim,
  epsilon,
  exclusion_region,
  genotype,
  hotspot,
  target_markers
)
print(sampling_set)

previous_sampling_set = sampling_set
set.seed(42)
exclusion_region = gatars_example$exclusion_region
head(exclusion_region)
sampling_set = gatars_sampling_set(
  bim,
  epsilon,
  exclusion_region,
  genotype,
  hotspot,
  target_markers
)
print(previous_sampling_set)
print(sampling_set)

# Calling gatars 
start = Sys.time()
set.seed(1)
gatars(phenotype, Psi, sampling_set, N_sim_reps = 100, weights = c(5, 3, 2))
elapsed_time = Sys.time() - start
paste("100 sim reps used", 
      round(elapsed_time, 1), attributes(elapsed_time)$units)

start = Sys.time()
set.seed(1)
gatars(phenotype, Psi, sampling_set, N_sim_reps = 1000, weights = c(5, 3, 2))
elapsed_time = Sys.time() - start
paste("1000 sim reps used", 
      round(elapsed_time, 1), attributes(elapsed_time)$units)

start = Sys.time()
set.seed(1)
gatars(phenotype, Psi, sampling_set, N_sim_reps = 2000, weights = c(5, 3, 2))
elapsed_time = Sys.time() - start
paste("2000 sim reps used", 
      round(elapsed_time, 1), attributes(elapsed_time)$units)

```

