---
title: "202-scribbling-with-gurobi"
author: ''
output:
  html_document:
    css: ~/Documents/headings.css
    fig_caption: yes
    highlight: default
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    pandoc_args: --variable=geometry:margin=0.75in
    toc: no
fontsize: 10pt
---
<A NAME="top"> </A>

```{r global_options, include=FALSE}
require(knitr)
opts_chunk$set(eval = TRUE, echo = FALSE, fig.height = 8, fig.width = 8, tidy=FALSE)
```

`r getwd()`  

Last knit was `r Sys.time()`

"2016-09-03 09:48:19 PDT"

# Preparing the data
```{r}
library(gritsr3)
library(gurobi)
library(Matrix)
bim = alternative_example$bim
genotype = alternative_example$genotype
fam = alternative_example$fam
target_markers = alternative_example$target_markers[3:5]
Psi = alternative_example$Psi
target_markers = target_markers

NNN = nrow(fam)
first_ten = 1:10
last_ten = NNN - (9:0)
matrix_image_fn(Psi[c(first_ten, last_ten), c(first_ten, last_ten)],
                main = "First and last 10 rows and columns of Psi")
```

# An example for `matrix_image_fn`

```{r}
NNN = 30
epsilon = .8
x = rnorm(NNN, 0, 1) 
x1 = x + rnorm(NNN, 0, epsilon)
x2 = x + rnorm(NNN, 0, epsilon)
x3 = x + rnorm(NNN, 0, epsilon)
independent_columns = matrix(rnorm(5 * NNN, 0, 1), ncol = 5)
data = cbind(x1, x2, x3, independent_columns)
matrix_image_fn(cor(data), "The correlation of some multivariate normal data")
```

# Checking the rank of the genotype_target_markers matrix

```{r}
library(Matrix)
genotype_target_markers = genotype[, target_markers]
list(target_markers = target_markers,
     rank = as.numeric(rankMatrix( genotype_target_markers)))
```

# This part contains my scribbling with gurobi

A call of `gritsr_sampling_set` consists of calling the following tow functions
```
  params_sampling_set = params_sampling_set_fn(
    bim,
    genotype,
    target_markers,
    hotspot,
    epsilon_on_log_scale,
    use_count_as_weight,
    use_popularity,
    use_expected,
    markers_needing_help,
    N_try)
  answer = sampling_set_fn(params_sampling_set)
```

```{r}
params_sampling_set = params_sampling_set_fn(
  bim,
  genotype,
  target_markers,
  hotspot,
  epsilon_on_log_scale = 0.02,
  use_count_as_weight = FALSE,
  use_popularity = FALSE,
  use_expected = FALSE,
  markers_needing_help = NULL,
  N_try = 100)
```

`sampling_set_fn` is defined
```

sampling_set_fn = function(params_sampling_set){
  genotype = params_sampling_set$genotype
  params_sampling_set$independent_segment = independent_segment =
    if(is.null(params_sampling_set$independent_segment)){
      independent_segment_fn(params_sampling_set)
    } else {
      params_sampling_set$independent_segment
    }
  if(is.null(params_sampling_set$assignment_probability) || is.null(params_sampling_set$the_matrix)) {
    aaa = assignment_probability_fn(params_sampling_set)
    params_sampling_set$assignment_probability = aaa$assignment_probability
    params_sampling_set$the_matrix = aaa$the_matrix
  }
  N_try = params_sampling_set$N_try
  p_target = params_sampling_set$p_target
  MMM = params_sampling_set$MMM
  epsilon_on_log_scale = params_sampling_set$epsilon_on_log_scale

  try = lapply(1:N_try, function(seed){
    params_sampling_set$seed = seed
    assign_segment_to_box_fn(params_sampling_set)
  })
  box = try[[which.max(sapply(try, function(this){
    min(this$sampling_set_size)
  }))]]$box

  sampling_set = lapply(1:MMM, function(mmm){
    this_box = sort(box[[mmm]])
    this_genotype = genotype[, do.call(rbind, independent_segment[this_box])$snp]
    p_sim = unname(colMeans(this_genotype)/2)
    p_this = p_target[mmm]
    lower = p_this * (1 - epsilon_on_log_scale)
    upper = p_this * (1 + epsilon_on_log_scale)
    good_ones_q = (lower < p_sim & p_sim < upper)
    answer =  this_genotype[, good_ones_q, drop = FALSE]
    answer
  })
  report = do.call(rbind, lapply(1:MMM, function(mmm){
    this_genotype = sampling_set[[mmm]]
    p_sim = unname(colMeans(this_genotype)/2)
    data.frame(min = round(min(p_sim), 5),
               p_target = round(p_target[mmm], 5),
               max = round(max(p_sim), 5),
               set_size = ncol(this_genotype))
  }))
  rownames(report) = NULL
  list(
    params_sampling_set = params_sampling_set,
    report = report,
    sampling_set = sampling_set)
}
```
```{r}
genotype = params_sampling_set$genotype
independent_segment_0 = independent_segment_fn(params_sampling_set)
aaa = assignment_probability_fn(params_sampling_set)
names(aaa)
the_matrix_0 = t(aaa$the_matrix)
keep = apply(the_matrix_0, 1, function(row){
  !all(row == 0)
})
the_matrix = the_matrix_0[keep, ]
independent_segment = independent_segment_0[keep]
list(dim_the_matrix = dim(the_matrix),
     length_independent_segment = length(independent_segment))
```


```{r}
if(FALSE){
keeper = apply(the_matrix, 1, function(row){
  sum(row != 0) >= 2
})
the_matrix = head(the_matrix[keeper, ], 7)
the_matrix
}
```

```{r}
source("MSparse-2016-09-02.R")

selection_matrix_fn = function(the_matrix){
  x_opt = MSparse(the_matrix)
  chosen = x_opt[1:(nrow(the_matrix) * ncol(the_matrix))]
  matrix(chosen, ncol = ncol(the_matrix), byrow = TRUE)
}
selection_matrix = selection_matrix_fn(the_matrix)
```
# box[[1]] tells me which of independent_segment to assign to sampling_set[1]
```{r}
box = apply(selection_matrix, 2, function(col){ 
  which(col == 1)
})

sampling_sizes = apply(selection_matrix * the_matrix, 2, sum)
```


```{r}
mmm = 1
p_target = params_sampling_set$p_target
epsilon_on_log_scale = params_sampling_set$epsilon_on_log_scale
MMM = params_sampling_set$MMM
sampling_set = lapply(1:MMM, function(mmm){
  this_box = box[[mmm]]
  this_genotype = genotype[, do.call(rbind, independent_segment[this_box])$snp]
  p_sim = unname(colMeans(this_genotype)/2)
  p_this = p_target[mmm]
  lower = p_this * (1 - epsilon_on_log_scale)
  upper = p_this * (1 + epsilon_on_log_scale)
  good_ones_q = (lower < p_sim & p_sim < upper)
  answer =  this_genotype[, good_ones_q, drop = FALSE]
  answer
})
```



# Creating the sampling_set
```{r}
set.seed(2)
sampling_set = gritsr_sampling_set(
  bim, genotype, target_markers, hotspot, 
  epsilon_on_log_scale = 0.02,
  use_count_as_weight = FALSE,
  use_popularity = FALSE,
  use_expected = FALSE,
  markers_needing_help = NULL,
  N_try = 100
)
print(sampling_set)

sampling_set = gritsr_sampling_set(
  bim, genotype, target_markers, hotspot, 
  epsilon_on_log_scale = 0.02,
  use_count_as_weight = FALSE,
  use_popularity = TRUE,
  use_expected = TRUE,
  markers_needing_help = c(2, 3),
  N_try = 100
)
print(sampling_set)
```



# Calling gritsr with N_sim_reps = 1000 and 2000
```{r}
start = Sys.time()
gritsr(fam, Psi, sampling_set, N_sim_reps = 1000)
elapsed_time = Sys.time() - start
paste("1000 sim reps required", 
      round(elapsed_time, 1), attributes(elapsed_time)$units)
start = Sys.time()
gritsr(fam, Psi, sampling_set, N_sim_reps = 2000)
elapsed_time = Sys.time() - start
paste("2000 sim reps required", 
      round(elapsed_time, 1), attributes(elapsed_time)$units)
```




```
#' print(sampling_set_for_gritsr_example)
#' N_sim_reps = 10
#' gritsr(fam, Psi, sampling_set_for_gritsr_example, N_sim_reps)
```