---
title: "03-params_sampling_set_fn"
author: ""
output:
  html_document:
    css: ~/Documents/headings.css
    fig_caption: yes
    highlight: default
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    pandoc_args: --variable=geometry:margin=0.75in
    toc: no
fontsize: 10pt
---
<A NAME="top"> </A>

```{r global_options, include=FALSE}
require(knitr)
opts_chunk$set(eval = TRUE, echo = TRUE, fig.height = 5, fig.width = 6, tidy=FALSE)
```

`r getwd()`  

"2017-09-14 08:12:36 PDT"

Last knit was `r Sys.time()`

```{r}
library(gatars)
bim = alternative_example$bim
epsilon_on_log_scale = 0.02
exclusion_region = data.frame(chromosome = integer(), begin = integer(), end = integer())
genotype = alternative_example$genotype
target_markers = alternative_example$target_markers[3:5]
params_sampling_set = params_sampling_set_fn(
  bim, epsilon_on_log_scale, exclusion_region,
  genotype, hotspot, target_markers)
names(params_sampling_set)
```

```{r}
names(params_sampling_set)
```

The following just come through as is:

`bim`, `genotype`, `target_markers`, `exclusion_region`, `hotspot`, `epsilon_on_log_scale`

`g_target` is a submatrix of `genotype` containing just the columns specified by `target_markers`.

`p_target` is a vector of length `MMM` whose `mmm`-th element is the mean of the `mmm`-th column of `g_target` divided by `2`.

`e_g_target` is a matrix with `NNN` rows and `MMM` columns.  All rows are the same.  The `mmm`-th column contains the mean of `g_target` at that column.

`MMM` is the length of `target_markers` and the number of columns of `g_target`.

`NNN` is the number of rows of `genotype` and `g_target`.



