---
title: "11-a-description-of-davies_not_lu_depends_on_g-target_fn"
author: ""
output:
  html_document:
    css: ~/Documents/headings.css
    fig_caption: yes
    highlight: default
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    pandoc_args: --variable=geometry:margin=0.75in
    toc: no
fontsize: 10pt
---
<A NAME="top"> </A>

```{r global_options, include=FALSE}
require(knitr)
opts_chunk$set(eval = TRUE, echo = TRUE, fig.height = 10, fig.width = 8, tidy=FALSE)
```

`r getwd()`  

"2017-09-25 19:09:11 PDT"

Last knit was `r Sys.time()`


# To get the distribution of the quadratic forms, we will use Davies, and therefore we need $\mu_Z$ and $V_Z$

# `davies_fn`

Given the $J$ dimensional random vector (of functions of genotypes) $Z \sim N \big( \mu_Z, V_Z \big)$ and the matrix $A$, 
the p-value for the statistic $Q = Z^T A Z$ can be gotten from `davies_fn(zzz, mu_z, V_z, AAA)`

**This function performs the following calculations**

$\tilde{Z} = V_{Z}^{-1/2} Z \sim N \big( \mu_{\tilde{Z}} = V_{Z}^{-1/2} \mu_Z, I \big)$.

$Q = Z^T A Z = Z^T V^{-1/2} V^{1/2} A V^{1/2} V^{-1/2} Z = \tilde{Z}^T \tilde{A} \tilde{Z}$

$\tilde{A} = V^{1/2} A V^{1/2}$

$\tilde{A} = U^T \Lambda U$. The spectral decomposition of $\tilde{A}$, so $\Lambda$ is a diagonal matrix containing the eigenvalues of $\tilde{A}$, and $U^T U = U U^T = I$.

$Q = \tilde{Z}^T \tilde{A} \tilde{Z} = \tilde{Z}^T U^T \Lambda U \tilde{Z} = X^T \Lambda X$

$X = U \tilde{Z} \sim N \big( \mu_X = U \mu_{\tilde{Z}}, U I U^T = I)$

$Q = X^T \Lambda X = \sum_{j = 1}^J \lambda_j X_j^2$ is a sum of independent $\chi^2$ random variables with noncentrality parameters $\mu_X^2$.

The R package `CompQuadForm` provides the function `davies` which computes $P(Q > q)$ where $Q = \sum_{j = 1}^J \lambda_j X_j + \sigma X_0$ where $X_j$ are independent randomvariables having a non-central $\chi^2$ distribution with $n_j$ degrees of freedom and noncentrality parameter $\delta_j^2$, and $X_0$ having a standard normal distribution.

In our case, the degrees of freedom is $n_j = 1$, and the noncentrality parameter is $\delta_j^2 = \mu_X^2$.

# The phenotype matrix

The phenoytpe matrix has `N` rows, one for each person, and the following columns

$Y$ is an indicator of disease or a continuous trait.  `y_1`

$\mu$ is a predictor for $y$.  `y_2`

$Psi$ is the kinship matrix (supplied by the user) $\Psi_{n_1, n_2} = \text{Cor}(G_{n_1, m}, G_{n_2, m})$ while holding $m$ equal to any number in `1, ..., M`.


# The genotype matrix

$G$  is the genotype matrix. `g_target` is an `N x M` matrix

$E(G)$. `e_g_target` is an `N x M` matrix

$\Gamma = cov(G_n)$. `V_G = cov(g_target)` is an `M x M` matrix

$W$ is an `M x M` diagonal matrix of weights.  These weights reflect relative importances of the $M$ target markers.

# The statistics we consider are quadratic forms $Z^T A Z$


$Z = \begin{pmatrix}
Z_1 \\
Z_2
\end{pmatrix} = \begin{pmatrix}
W G Y \\
W G \mu
\end{pmatrix}$

$Z_k = 1(k \le M) \sum_{n = 1}^N w_k g_{k n} y_n + 1(k > M) \sum_{n = 1}^N w_{k-M+1} g_{k-M+1} \mu_n$


Each term $\text{cov}(Z_{k_1}, Z_{k_2})$ is of the form (in the following expression I need to replace $k_1$ with $k_1\text{mod}K$
and $k_2$ with $k_2\text{mod}K$) 
$\text{cov}\Big( \sum_{n_1 = 1}^N w_{k_1} g_{k_1 n_1} a_{n_1}, \sum_{n_2 = 1}^N w_{k_2} g_{k_2 n_2} b_{n_2} \Big)$

Where $a_{n_1}$ and $b_{n_2}$ are either $y_{n_1}$ or $\mu_{n_2}$ depending upon whether each of $k_1$ and $k_2$ do not or do exceed $M$.


$\text{cov}\Big( \sum_{n_1 = 1}^N w_{k_1} g_{k_1 n_1} a_{n_1}, \sum_{n_2 = 1}^N w_{k_2} g_{k_2 n_2} b_{n_2} \Big)
= w_{k_1} w_{k_2} \sum_{n_1 = 1}^N   a_{n_1} \sum_{n_2 = 1}^N  b_{n_2}\text{cov}\Big(g_{k_1 n_1}, g_{k_2 n_2} \Big)$
$= w_{k_1} w_{k_2} \sum_{n_1 = 1}^N   a_{n_1} \sum_{n_2 = 1}^N  b_{n_2} \Gamma_{k_1, k_2} \Psi_{n_1, n_2}
= w_{k_1}\Gamma_{k_1, k_2} w_{k_2} \sum_{n_1 = 1}^N   \sum_{n_2 = 1}^N a_{n_1}  \Psi_{n_1, n_2} b_{n_2}$
$= \Big(W \Gamma W \Big)_{k_1, k_2} a^T \Psi b = a^T \Psi b \Big(W \Gamma W \Big)_{k_1, k_2}$

$\text{cov}(Z) = \begin{pmatrix}
y^T   \Psi y \;  W \Gamma W  & y^T   \Psi \mu \; W \Gamma W   \\
\mu^T \Psi y \;  W \Gamma W  & \mu^T \Psi \mu \; W \Gamma W  
\end{pmatrix}$


The R function `kronecker` behaves like this

```{r echo = TRUE}
a = matrix(1:6, nrow = 3)
b = matrix((1:4) * 10, nrow = 1)
list(a = a,
     b = b,
     kronecker = kronecker(a, b))
```

And so I believe I want to write

$\text{cov} = \text{kronecker} \Big(  \begin{pmatrix}
 y^T \Psi y   &   y^T \Psi \mu \\
 \mu^T \Psi y &  \mu^T \Psi \mu
\end{pmatrix} \otimes W \Gamma W 
\Big) = \text{kronecker}  \begin{pmatrix}
y^T \\
\mu^T
\end{pmatrix} \Psi \begin{pmatrix} y & \mu \end{pmatrix} \otimes W \Gamma W$

The first argument in the the `kronecker` function 
$\begin{pmatrix}
y^T \\
\mu^T
\end{pmatrix} \Psi \begin{pmatrix} y & \mu \end{pmatrix}$ is a $2 \times 2$ matrix.
The second argument $W \Gamma W$ is an $M \times M$ matrix, and so the `kronecker` of the two is a $2M \times 2M$ matrix.






