---
title: "19-for-now-just-correct-davies_not_lu"
author: ""
output:
  html_document:
    css: ~/Documents/headings.css
    fig_caption: yes
    highlight: default
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    pandoc_args: --variable=geometry:margin=0.75in
    toc: no
fontsize: 10pt
---
<A NAME="top"> </A>

```{r global_options, include=FALSE}
require(knitr)
opts_chunk$set(eval = TRUE, echo = FALSE, fig.height = 10, fig.width = 8, tidy=FALSE)
```

`r getwd()`  

"2017-09-26 11:06:13 PDT"

Last knit was `r Sys.time()`

```{r}
library(gatars)
keeper = c(1:12, 101:112)
bim = alternative_example$bim
genotype = alternative_example$genotype[keeper, ]
fam = alternative_example$fam[keeper, ]
target_markers = alternative_example$target_markers[c(3, 5)]
Psi = alternative_example$Psi[keeper, keeper]
exclusion_region = alternative_example$exclusion_region
genotype_target_markers = genotype[, target_markers]
set.seed(2)
epsilon_on_log_scale = 0.2
sampling_set = gatars_sampling_set(
    bim, epsilon_on_log_scale, exclusion_region,
    genotype, hotspot, target_markers)
print(sampling_set)

N_sim_reps = 10
# gatars(fam, Psi, sampling_set, N_sim_reps = 10)
params_sampling_set = sampling_set$params_sampling_set
sampling_set = sampling_set$sampling_set
weights = NULL
params = params_fn(params_sampling_set, fam, Psi, sampling_set, N_sim_reps, weights)
# ooo = one_experiment_fn(params, calculate_optimized = TRUE)
# ooo$p_value
# inside one_experiment_fn
calculate_optimized = TRUE
theta_init = params$theta_init
g_target = params$g_target
library(Matrix)
rankMatrix(g_target)
# ooo = one_experiment_fn(params, calculate_fancy = TRUE)
# ooo
```


```{r}
g_target = params$g_target
MMM = params$MMM
WWW = params$WWW
www = params$www
y_1 = params$yyy
y_2 = params$e_y
rho_uni = params$rho_uni
Phi = params$Phi
```

# `davies_not_lu_depends_on_g_target_fn`
```{r}
JJJ = matrix(rep(1, MMM * MMM), nrow = MMM)
III = diag(rep(1, MMM))
zero_M = matrix(rep(0, MMM * MMM), nrow = MMM)
zero_M_2M = matrix(rep(0, MMM * 2 * MMM), nrow = MMM)
z_1 = as.vector(WWW %*% t(g_target) %*% y_1)
z_2 = as.vector(WWW %*% t(g_target) %*% y_2)
zzz = c(z_1, z_2)
e_g_target_1 = colMeans(g_target)
e_g_target = matrix(rep(e_g_target_1, nrow(g_target)), nrow = nrow(g_target), byrow = TRUE)
e_z_1 = matrix(as.vector(WWW %*% t(e_g_target) %*% y_1), ncol = 1)
e_z_2 = matrix(as.vector(WWW %*% t(e_g_target) %*% y_2), ncol = 1)
mu_z = rbind(e_z_1, e_z_2)
V_G = cov(g_target)
W_VG_W = WWW %*% V_G %*% WWW
V_z = kronecker(Phi, W_VG_W)
# ----------------------------------------------- straight, fancy_grid
alpha_uni_zero = expand.grid(a1 = rho_uni, a2 = rho_uni)
alpha_uni = alpha_uni_zero[apply(alpha_uni_zero, 1, sum) <= 1, ]
rownames(alpha_uni) = NULL
str(alpha_uni)
```


```{r}
AAA_fn = function(alpha_B, alpha_S, alpha_T, MMM){
  III = diag(rep(1, MMM))
  JJJ = matrix(rep(1, MMM * MMM), nrow = MMM)
  A_1 = alpha_B * JJJ + (alpha_S + alpha_T) * III
  A_2 = -(alpha_B * JJJ + alpha_S * III)
  A_3 = alpha_B * JJJ + (alpha_S - alpha_T) * III
  AAA = rbind(cbind(A_1, A_2),
              cbind(A_2, A_3))
  AAA
}
zzz_etc_fn = function(g_target, Phi, WWW, y_1, y_2){
  z_1 = as.vector(WWW %*% t(g_target) %*% y_1)
  z_2 = as.vector(WWW %*% t(g_target) %*% y_2)
  zzz = c(z_1, z_2)
  
  e_g_target_1 = colMeans(g_target)
  e_g_target = matrix(rep(e_g_target_1, nrow(g_target)), nrow = nrow(g_target), byrow = TRUE)
  e_z_1 = matrix(as.vector(WWW %*% t(e_g_target) %*% y_1), ncol = 1)
  e_z_2 = matrix(as.vector(WWW %*% t(e_g_target) %*% y_2), ncol = 1)
  mu_z = rbind(e_z_1, e_z_2)
  
  V_G = cov(g_target)
  W_VG_W = WWW %*% V_G %*% WWW
  V_z = kronecker(Phi, W_VG_W)
  
  list(zzz = zzz,
       mu_z = mu_z,
       V_z = V_z)
}
```

```{r echo = TRUE}
rho_uni_times_10 = 0:10
alpha_uni_zero = expand.grid(
  B = rho_uni_times_10,
  S = rho_uni_times_10,
  T = rho_uni_times_10)
str(alpha_uni_zero)
alpha_uni_zero_sum = apply(alpha_uni_zero, 1, sum)
alpha_uni = alpha_uni_zero[alpha_uni_zero_sum == 10, ]
row.names(alpha_uni) = NULL
alpha_uni = alpha_uni/10
```
# ---------------------------------------------------- zzz etc
zzz_etc = zzz_etc_fn(g_target, Phi, WWW, y_1, y_2)
zzz = zzz_etc$zzz
mu_z = zzz_etc$mu_z
V_z = zzz_etc$V_z
# -------------------------------------------------- alpha_uni
alpha_uni_zero = expand.grid(
  B = rho_uni, 
  S = rho_uni, 
  T = rho_uni)
alpha_uni_zero_sum = apply(alpha_uni_zero, 1, sum)
alpha_uni = alpha_uni_zero[alpha_uni_zero_sum == 1, ]
row.names(alpha_uni) = NULL


another_0 = expand.grid(
  B = rho_uni,
  S = rho_uni)
another = another_0[apply(another_0, 1, sum) <= 1, ]
row.names(another) = NULL
another$T = round(1 - another$B - another$S, 1)

another[do.call(order, another), ]
alpha_uni[do.call(order, alpha_uni), ]
```

