---
title: "07-basic_and_optimized_lu_fn"
author: ""
output:
  html_document:
    css: ~/Documents/headings.css
    fig_caption: yes
    highlight: default
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    pandoc_args: --variable=geometry:margin=0.75in
    toc: no
fontsize: 10pt
---
<A NAME="top"> </A>

```{r global_options, include=FALSE}
require(knitr)
opts_chunk$set(eval = TRUE, echo = TRUE, fig.height = 10, fig.width = 8, tidy=FALSE)
```

`r getwd()`  

"2017-11-17 09:07:47 PST"

Last knit was `r Sys.time()`

```{r}
library(gatars)
# Preparing the data
bim = gatars_example$bim
genotype = gatars_example$genotype
phenotype = gatars_example$phenotype
Psi = gatars_example$Psi
target_markers = gatars_example$target_markers[3:5]

# Call gatars_sampling_set to create sampling_set
sampling_set = gatars_sampling_set(
  bim, epsilon = 0.01, exclusion_region = NULL,
  genotype, hotspot, target_markers)
print(sampling_set)

# Call gatars_test_size using an N_simulated_nulls which is way too small
N_simulated_nulls = 10
library(Matrix)
gatars_test_size(phenotype, Psi, sampling_set, N_simulated_nulls)
```

# inside `gatars_test_size`
```{r}
params_sampling_set = sampling_set$params_sampling_set
sampling_set = sampling_set$sampling_set
alpha_uni_N_increments = 10
weights = NULL
params = params_fn(
    alpha_uni_N_increments, params_sampling_set, phenotype, Psi, sampling_set, N_simulated_nulls, weights)
```


```{r}
ooo = uno_experimento_fn(params, calculate_optimized = TRUE)
```

# inside `uno_experimento_fn`

```{r}
calculate_optimized = TRUE
adaptive_conf_level = params$adaptive_conf_level
alpha_uni = params$alpha_uni
g_target = params$g_target
MMM = params$MMM
N_simulated_nulls_interval = params$N_simulated_nulls_interval
N_simulated_nulls_limit = params$N_simulated_nulls_limit
theta_init = params$theta_init
WWW = params$WWW
y_1 = params$yyy
y_2 = params$mu
Phi = params$Phi
sampling_set = params$sampling_set
test_size = params$test_size
theta_init = params$theta_init
bo = basic_and_optimized_lu_fn(alpha_uni, g_target, Phi, theta_init, WWW, y_1, y_2)
bo
```


# inside `basic_and_optimized_lu_fn = function(alpha_uni, g_target, Phi, theta, WWW, y_1, y_2){`
```{r}
# ---------------------------------------------------- zzz etc
zzz_etc = zzz_etc_fn(g_target, Phi, WWW, y_1, y_2)
zzz_etc
```

# insdie `z_etc_fn`
In the notation of the web site, 

`y_1` is `phenotype$y` or $y$ the $N \times 1$ column vector of subjects' coded trait phenotypes.     
`y_2` is `phenotype$mu`or $\mu$ the $N \times 1$ column vector of user-specified phenotype predictions.    
`WWW` is $W$ the diagonal matrix whose user-specifed diagonal components weight the target markers.     

`Phi` described in `05-2015-10-18-unified-kernel-statistcs-add-other-optimizations.pdf`. In the manuscript, `V_z = kronecker(Phi, W_VG_W)` where 
`Phi_{k_1, k_2} = y_{k_1} Psi y_k_2` = 

```
Phi_fn = function(Psi, y_1, y_2){
  Phi_11 = t(y_1) %*% Psi %*% y_1
  Phi_12 = t(y_1) %*% Psi %*% y_2
  Phi_22 = t(y_2) %*% Psi %*% y_2
  Phi = matrix(c(Phi_11, Phi_12, Phi_12, Phi_22), nrow = 2)
  Phi
}
```

Returns:
`zzz =` $Z$ where `zzz = c(z_1, z_2)` and `z_1 = ` $W G y$ and `z_2 = ` $W G \mu$     
`mu_z` where `mu_z_1 = ` $W E(G) y$ and `mu_z_2 =` $W E(G) \mu$     
`V_z` the covariance matrix of `zzz`. `V_z = kronecker(Phi, W_VG_W)`

```{r}
WWW
Phi
z_1 = as.vector(WWW %*% t(g_target) %*% y_1)
z_2 = as.vector(WWW %*% t(g_target) %*% y_2)
zzz = c(z_1, z_2)

e_g_target_1 = colMeans(g_target)
e_g_target = matrix(rep(e_g_target_1, nrow(g_target)), nrow = nrow(g_target), byrow = TRUE)
e_z_1 = matrix(as.vector(WWW %*% t(e_g_target) %*% y_1), ncol = 1)
e_z_2 = matrix(as.vector(WWW %*% t(e_g_target) %*% y_2), ncol = 1)
mu_z = rbind(e_z_1, e_z_2)

V_G = cov(g_target)
W_VG_W = WWW %*% V_G %*% WWW
V_z = kronecker(Phi, W_VG_W)

list(zzz = zzz,
     mu_z = mu_z,
     V_z = V_z)
```

