---
title: "07-basic_and_optimized_lu_fn"
author: ""
output:
  html_document:
    css: ~/Documents/headings.css
    fig_caption: yes
    highlight: default
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    pandoc_args: --variable=geometry:margin=0.75in
    toc: no
fontsize: 10pt
---
<A NAME="top"> </A>

```{r global_options, include=FALSE}
require(knitr)
opts_chunk$set(eval = TRUE, echo = TRUE, fig.height = 10, fig.width = 8, tidy=FALSE)
```

`r getwd()`  

"2017-11-17 09:07:47 PST"

Last knit was `r Sys.time()`


# Calculate $Z$ and its first two moments.

In the manuscript, the first two moments are denoted $E_0(Z)$ and $Cov_0(Z)$ and are given by equation (8).

In `gatars, the first two moments are denoted `mu_z` and `V_z`.


```{r}
library(gatars)
# Preparing the data
bim = gatars_example$bim
genotype = gatars_example$genotype
phenotype = gatars_example$phenotype
Psi = gatars_example$Psi
target_markers = gatars_example$target_markers[3:5]

g_target = genotype[, target_markers]

y_1 = yyy = phenotype$y
y_2 = mu = phenotype$mu

Phi = Phi_fn(Psi, y_1, y_2)
Phi
MMM = length(target_markers)
WWW =  diag(rep(1, MMM))

zzz_etc = zzz_and_first_two_moments_fn(g_target, Phi, WWW, y_1, y_2)
zzz_etc
str(zzz_etc)
```


`z_etc_fn` should be renamed `z_and_first_two_moments_fn`

In the notation of the web site, 

`g_target` is the genotype matrix $G$ in the manuscript.

`y_1` is `phenotype$y` or $y$ the $N \times 1$ column vector of subjects' coded trait phenotypes.      

`y_2` is `phenotype$mu`or $\mu$ the $N \times 1$ column vector of user-specified phenotype predictions.   

`WWW` is $W$ the diagonal matrix whose user-specifed diagonal components weight the target markers.     

`Phi` `Phi_{k_1, k_2} = y_{k_1} Psi y_k_2`.  In the manuscript, equation 8,  `V_z = kronecker(Phi, W_VG_W)`.

```
Phi_fn = function(Psi, y_1, y_2){
  Phi_11 = t(y_1) %*% Psi %*% y_1
  Phi_12 = t(y_1) %*% Psi %*% y_2
  Phi_22 = t(y_2) %*% Psi %*% y_2
  Phi = matrix(c(Phi_11, Phi_12, Phi_12, Phi_22), nrow = 2)
  Phi
}
```

Returns:
`zzz =` $Z$ where `zzz = c(z_1, z_2)` and `z_1 = ` $W G y$ and `z_2 = ` $W G \mu$      
`mu_z` where `mu_z_1 = ` $W E(G) y$ and `mu_z_2 =` $W E(G) \mu$ (see equation 8 of the manuscript).     
`V_z` the covariance matrix of `zzz`. `V_z = kronecker(Phi, W_VG_W)` (see equation 8 of the manuscript).



