---
title: "11-davies_fn"
author: ""
output:
  html_document:
    css: ~/Documents/headings.css
    fig_caption: yes
    highlight: default
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    pandoc_args: --variable=geometry:margin=0.75in
    toc: no
fontsize: 10pt
---
<A NAME="top"> </A>

```{r global_options, include=FALSE}
require(knitr)
opts_chunk$set(eval = TRUE, echo = TRUE, fig.height = 10, fig.width = 8, tidy=FALSE)
```

`r getwd()`  

"2017-11-20 07:58:11 PST"

I now using gatars v-0.2.17

Last knit was `r Sys.time()`

```{r}
library(gatars)
bim = gatars_example$bim
genotype = gatars_example$genotype
phenotype = gatars_example$phenotype
Psi = gatars_example$Psi
target_markers = gatars_example$target_markers[3:5]
g_target = genotype[, target_markers]
MMM = ncol(g_target)
NNN = nrow(g_target)
e_g_target_1 = colMeans(g_target)
p_target = e_g_target_1/2
e_g_target = matrix(rep(e_g_target_1, nrow(g_target)), nrow = nrow(g_target), byrow = TRUE)
y_1 = yyy = phenotype$y
y_2 = mu = phenotype$mu
Phi = Phi_fn(Psi, y_1, y_2)
www_num = rep(1, MMM)
www = www_num/sum(www_num) * MMM
WWW = diag(www)
theta_init = rep(pi/3, 3)
bo = basic_and_optimized_lu_fn(g_target, Phi, theta_init, WWW, y_1, y_2)
bo
```


# inside `basic_and_optimized_lu_fn = function(g_target, Phi, theta, WWW, y_1, y_2){`
```{r}
# ---------------------------------------------------- zzz etc
zzz_etc = zzz_and_first_two_moments_fn(g_target, Phi, WWW, y_1, y_2)
zzz = zzz_etc$zzz
mu_z = zzz_etc$mu_z
V_z = zzz_etc$V_z
AAA = AAA_fn(1, 0, 0, MMM)
davies_fn(zzz, mu_z, V_z, AAA)
```


`AAA_fn` calculates $A_{\alpha}$ in equation (7) in the manuscript

`alpha_B`, `alpha_S`, and `alpha_T` are real numbers in the closed interval
`[0, 1], and they add to `1`.

`MMM` is an integer equal to the length of `target_markers`.     

`AAA_fn` returns a matrix of dimension `(2 * MMM)` by `(2 * MMM)`.

```
AAA_fn = function(alpha_B, alpha_S, alpha_T, MMM){
  III = diag(rep(1, MMM))
  JJJ = matrix(rep(1, MMM * MMM), nrow = MMM)
  A_1 = alpha_B * JJJ + (alpha_S + alpha_T) * III
  A_2 = -(alpha_B * JJJ + alpha_S * III)
  A_3 = alpha_B * JJJ + (alpha_S - alpha_T) * III
  AAA = rbind(cbind(A_1, A_2),
              cbind(A_2, A_3))
  AAA
}
```

`davies_fn` calculates the spectral decomposition of $Q = Z^T A Z$ to obtain
the eigenvalues $\lambda_j, j = 1, \cdots, J$ and noncentrality parameter
$\delta$ that are required by `davies` in the `CompquadForm` package.

```
#' @import CompQuadForm
#' @export
davies_fn = function(zzz, mu_z, V_z, AAA){
  qqq = as.vector(t(zzz) %*% AAA %*% zzz)
  sss = square_root_matrix_fn(V_z)
  V_z_one_half = sss$A_one_half
  V_z_minus_one_half = sss$A_minus_one_half
  z_tilde = V_z_minus_one_half %*% zzz
  mu_z_tilde =  V_z_minus_one_half %*% mu_z
  V_z_tilde = V_z_minus_one_half %*% V_z %*% V_z_minus_one_half
  A_tilde = V_z_one_half %*% AAA %*% V_z_one_half
  # ------------------------------------------------- breve, lambda, delta
  eigen_A_tilde = eigen(A_tilde)
  UUU = Re(t(eigen_A_tilde$vectors))
  lambda = Re(eigen_A_tilde$values)
  Lambda = diag(lambda)
  z_breve = UUU %*% z_tilde
  mu_z_breve = UUU %*% mu_z_tilde
  delta = mu_z_breve^2
  q_spectral_decomp = sum(lambda * z_breve^2)
  # ----------------------------------------------- p_value
  p_value = davies(q = qqq, lambda = lambda, delta = delta,
                   lim = 50000, acc = 0.00005)$Qq
  data.frame(q = qqq, p_value)
}
```

I copied the following from 
`/Volumes/lacie/aaa/packages/gritsr2/gails-stuff/b-examples/02-v-9010/a-docum/`
`05-2015-10-18-unified-kernel-statistcs-add-other-optimizations.Rmd`

# `davies_fn`

Given the $J$ dimensional random vector (of functions of genotypes) $Z \sim N \big( \mu_Z, V_Z \big)$ and the matrix $A$, 
the p-value for the statistic $Q = Z^T A Z$ can be gotten from `davies_fn(zzz, mu_z, V_z, AAA)`

**This function performs the following calculations**

$\tilde{Z} = V_{Z}^{-1/2} Z \sim N \big( \mu_{\tilde{Z}} = V_{Z}^{-1/2} \mu_Z, I \big)$.

$Q = Z^T A Z = Z^T V^{-1/2} V^{1/2} A V^{1/2} V^{-1/2} Z = \tilde{Z}^T \tilde{A} \tilde{Z}$

$\tilde{A} = V^{1/2} A V^{1/2}$

$\tilde{A} = U^T \Lambda U$. The spectral decomposition of $\tilde{A}$, so $\Lambda$ is a diagonal matrix containing the eigenvalues of $\tilde{A}$, and $U^T U = U U^T = I$.

$Q = \tilde{Z}^T \tilde{A} \tilde{Z} = \tilde{Z}^T U^T \Lambda U \tilde{Z} = X^T \Lambda X$

$X = U \tilde{Z} \sim N \big( \mu_X = U \mu_{\tilde{Z}}, U I U^T = I)$

$Q = X^T \Lambda X = \sum_{j = 1}^J \lambda_j X_j^2$ is a sum of independent $\chi^2$ random variables with noncentrality parameters $\mu_X^2$.

The R package `CompQuadForm` provides the function `davies` which computes $P(Q > q)$ where $Q = \sum_{j = 1}^J \lambda_j X_j + \sigma X_0$ where $X_j$ are independent randomvariables having a non-central $\chi^2$ distribution with $n_j$ degrees of freedom and noncentrality parameter $\delta_j^2$, and $X_0$ having a standard normal distribution.

In our case, the degrees of freedom is $n_j = 1$, and the noncentrality parameter is $\delta_j^2 = \mu_X^2$.
