---
title: "15-genome_resampling_fn"
author: ""
output:
  html_document:
    css: ~/Documents/headings.css
    fig_caption: yes
    highlight: default
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    pandoc_args: --variable=geometry:margin=0.75in
    toc: no
fontsize: 10pt
---
<A NAME="top"> </A>

```{r global_options, include=FALSE}
require(knitr)
opts_chunk$set(eval = TRUE, echo = TRUE, fig.height = 10, fig.width = 8, tidy=FALSE)
```

`r getwd()`  

"2017-11-20 15:50:11 PST"

I now using gatars v-0.2.17

Last knit was `r Sys.time()`

```{r}
library(gatars)
library(Matrix)
bim = gatars_example$bim
genotype = gatars_example$genotype
phenotype = gatars_example$phenotype
Psi = gatars_example$Psi
target_markers = gatars_example$target_markers[3:5]
g_target = genotype[, target_markers]
MMM = ncol(g_target)
NNN = nrow(g_target)
e_g_target_1 = colMeans(g_target)
p_target = e_g_target_1/2
e_g_target = matrix(rep(e_g_target_1, nrow(g_target)), nrow = nrow(g_target), byrow = TRUE)
y_1 = yyy = phenotype$y
y_2 = mu = phenotype$mu
Phi = Phi_fn(Psi, y_1, y_2)
www_num = rep(1, MMM)
www = www_num/sum(www_num) * MMM
WWW = diag(www)
zzz_etc = zzz_and_first_two_moments_fn(g_target, Phi, WWW, y_1, y_2)
zzz = zzz_etc$zzz
mu_z = zzz_etc$mu_z
V_z = zzz_etc$V_z
AAA = AAA_fn(1, 0, 0, MMM)
theta_init = rep(pi/3, 2)
bo = basic_and_optimized_lu_fn(g_target, Phi, theta_init, WWW, y_1, y_2)
bo$xxx
bo$theta
theta = bo$theta
x_observed = bo$xxx
adaptive_conf_level = 0.01
calculate_optimized = TRUE
test_size = 0.01
epsilon = 0.01
exclusion_region = NULL
sampling_set = gatars_sampling_set(
  bim, epsilon, exclusion_region,
  genotype, hotspot, target_markers)
print(sampling_set)
sampling_set = sampling_set$sampling_set
str(sampling_set)
set.seed(1)
N_simulated_nulls_interval = 7
optimized_names = names(x_observed)
sss = genome_resampling_fn(MMM, N_simulated_nulls_interval, optimized_names, Phi, 
                                 sampling_set, theta, WWW, y_1, y_2)
sss
str(sss)
```

Accumulate a matrix with `N_simulated_nulls_interval` rows and a column for each of the optimized statistics.
Each row contains the optimized statistics for one simulated genotype matrix.
The reason this function looks so complicated is there are two `while` loops.
The inside loop generates a `genotype_sim` and then checks to see if it has full rank.
If not it tries again; there is a limit of 1000 tries.
The outside loop accumulates `N_simulated_nulls_interval` good tries.

`so_far_so_good` is a logical equal to `TRUE` if the number of bad tries never exceeded 1000 in all 
`N_simulated_nulls_interval` loops.

# inside `genome_resampling`

```{r}
set.seed(1)
genotype_sim = do.call(cbind, lapply(sampling_set, function(this_snp){
  this_snp[, sample(1:ncol(this_snp), 1)]
}))
rankMatrix(genotype_sim)
bo = basic_and_optimized_lu_fn(genotype_sim, Phi, theta, WWW, y_1, y_2)
one_row_in_simulated = bo$xxx
one_row_in_simulated
```



```{r}
# genome_resampling_fn = function(
#   MMM,
#   N_simulated_nulls_interval,
#   optimized_names,
#   Phi,
#   sampling_set,
#   theta, 
#   WWW,
#   y_1,
#   y_2
# ){
  simulated = data.frame(rep(NA, N_simulated_nulls_interval), 
                         rep(NA, N_simulated_nulls_interval), 
                         rep(NA, N_simulated_nulls_interval), 
                         rep(NA, N_simulated_nulls_interval))
  names(simulated) = optimized_names
  so_far_so_good = TRUE
  n_sim = 1
  while(so_far_so_good & n_sim <= N_simulated_nulls_interval){
    still_looking_for_genotype_sim = TRUE
    N_bad_genotype_sim = 0
    while(still_looking_for_genotype_sim){
      genotype_sim = do.call(cbind, lapply(sampling_set, function(this_snp){
        this_snp[, sample(1:ncol(this_snp), 1)]
      }))
      if(rankMatrix(genotype_sim) == MMM) {
        good_genotype_sim = TRUE
        still_looking_for_genotype_sim = FALSE
      } else {
        print("genotype_sim not full rank")
        N_bad_genotype_sim = N_bad_genotype_sim + 1
        if(N_bad_genotype_sim > 1000) {
          good_genotype_sim = FALSE
          still_looking_for_genotype_sim = FALSE
        }
      }
    }
    if(good_genotype_sim){
      bo = basic_and_optimized_lu_fn(genotype_sim, Phi, theta, WWW, y_1, y_2)
      one_row_in_simulated = bo$xxx
      simulated[n_sim, ] = one_row_in_simulated
      n_sim = n_sim + 1
    } else {
      so_far_so_good = FALSE
    }
  }
  list(
    simulated = simulated,
    so_far_so_good = so_far_so_good
  )
# }
```

