---
title: "16-rejuvenate_successes_fn"
author: ""
output:
  html_document:
    css: ~/Documents/headings.css
    fig_caption: yes
    highlight: default
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    highlight: tango
    number_sections: yes
    pandoc_args: --variable=geometry:margin=0.75in
    toc: no
fontsize: 10pt
---
<A NAME="top"> </A>

```{r global_options, include=FALSE}
require(knitr)
opts_chunk$set(eval = TRUE, echo = TRUE, fig.height = 10, fig.width = 8, tidy=FALSE)
```

`r getwd()`  

"2017-11-20 15:50:11 PST"

I now using gatars v-0.2.17

Last knit was `r Sys.time()`

```{r}
library(gatars)
library(Matrix)
bim = gatars_example$bim
genotype = gatars_example$genotype
phenotype = gatars_example$phenotype
Psi = gatars_example$Psi
target_markers = gatars_example$target_markers[3:5]
g_target = genotype[, target_markers]
MMM = ncol(g_target)
NNN = nrow(g_target)
e_g_target_1 = colMeans(g_target)
p_target = e_g_target_1/2
e_g_target = matrix(rep(e_g_target_1, nrow(g_target)), nrow = nrow(g_target), byrow = TRUE)
y_1 = yyy = phenotype$y
y_2 = mu = phenotype$mu
Phi = Phi_fn(Psi, y_1, y_2)
www_num = rep(1, MMM)
www = www_num/sum(www_num) * MMM
WWW = diag(www)
zzz_etc = zzz_and_first_two_moments_fn(g_target, Phi, WWW, y_1, y_2)
zzz = zzz_etc$zzz
mu_z = zzz_etc$mu_z
V_z = zzz_etc$V_z
AAA = AAA_fn(1, 0, 0, MMM)
theta_init = rep(pi/3, 2)
bo = basic_and_optimized_lu_fn(g_target, Phi, theta_init, WWW, y_1, y_2)
bo$xxx
bo$theta
theta = bo$theta
x_observed = bo$xxx
adaptive_conf_level = 0.01
calculate_optimized = TRUE
test_size = 0.01
epsilon = 0.01
exclusion_region = NULL
sampling_set = gatars_sampling_set(
  bim, epsilon, exclusion_region,
  genotype, hotspot, target_markers)
print(sampling_set)
sampling_set = sampling_set$sampling_set
str(sampling_set)
set.seed(1)
N_simulated_nulls_limit = 100
N_simulated_nulls_interval = 20
optimized_names = names(x_observed)

successes = rep(0, length(x_observed))
names(successes) = names(x_observed)
N_simulated_nulls_required = 0
still_looking = TRUE
# inside a while loop that continues until still_looking is set to FALSE
sss = genome_resampling_fn(MMM, N_simulated_nulls_interval, optimized_names, Phi, 
                           sampling_set, theta, WWW, y_1, y_2)
simulated = sss$simulated
simulated
x_observed
x_observed["BST"] < simulated$BST
so_far_so_good = sss$so_far_so_good

uuu = rejuvenate_successes_fn(adaptive_conf_level, N_simulated_nulls_limit, N_simulated_nulls_required,
                              optimized_names, simulated, so_far_so_good, successes, test_size, x_observed)
uuu
N_simulated_nulls_required = uuu$N_simulated_nulls_required
successes = uuu$successes
still_looking = uuu$still_looking
```


# inside `rejuvenate_successes_fn



```{r}
    more = unlist(sapply(optimized_names, function(this){
      sum(simulated[[this]] > x_observed[[this]])
    }, simplify = FALSE))
    successes = successes + more
    N_simulated_nulls_required = N_simulated_nulls_required + nrow(simulated)
    inside_ambiguous =  lapply(successes, function(xxx){
      ci = prop.test(xxx,
                     N_simulated_nulls_required,
                     conf.level = adaptive_conf_level)$conf.int
    })
    inside_ambiguous
    ambiguous = any(sapply(successes, function(xxx){
      ci = prop.test(xxx,
                     N_simulated_nulls_required,
                     conf.level = adaptive_conf_level)$conf.int
      ci[1] <= .10
    }))
    still_looking = ambiguous && N_simulated_nulls_required < N_simulated_nulls_limit
```

```
# rejuvenate_successes_fn = function(
#   adaptive_conf_level,
#   N_simulated_nulls_limit,
#   N_simulated_nulls_required,
#   optimized_names,
#   simulated,
#   so_far_so_good,
#   successes,
#   test_size,
#   x_observed
# ){
  if(so_far_so_good){
    more = unlist(sapply(optimized_names, function(this){
      sum(simulated[[this]] > x_observed[[this]])
    }, simplify = FALSE))
    successes = successes + more
    N_simulated_nulls_required = N_simulated_nulls_required + nrow(simulated)
    ambiguous = any(sapply(successes, function(xxx){
      ci = prop.test(xxx,
                     N_simulated_nulls_required,
                     conf.level = adaptive_conf_level)$conf.int
      ci[1] <= .10
    }))
    still_looking = ambiguous && N_simulated_nulls_required < N_simulated_nulls_limit
  } else { # prepare to jetison
    still_looking = FALSE
  }
  list(
    N_simulated_nulls_required = N_simulated_nulls_required,
    successes = successes,
    still_looking = still_looking)
# }
```
`p_value_optimized_fn` estimates the p-values of the optimized statistics.  At its most basic level, this function generates a "large number" of simulated null statistics to compare to the observed. The p-value is the proportion of these simulated nulls that exceed the observed.  The user can specify how big to make "a large number". For the simulations, I added another level of control:  If after generating a moderate number of simulated nulls, it becomes obvious that the observed of all four optimized statistics is not going to be significant, there is no longer any need to continue to the big number that might be needed if at least one of the significance levels were close to the desired significance level.

To specify "a very large number", use `N_simulated_nulls_limit`.
To allow `p_value_optimized_fn` to abort when it becomes obvious that all four optimized statistics will not be significant,
use `N_simulated_nulls_interval` to specify a moderate number of simulated nulls to calculate to get an idea of the significance level of each statistic.
`gatars_test_size` allows the user to set `N_simulated_nulls` to be the very large number and then sets
`N_simulated_nulls_interval = N_simulated_nulls_limit = N_simulated_nulls`.


v-0.2.19
I removed test_size from
p_value_optimized_fn
rejuvenate_successes_fn
uno_experimento_fn